---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# spotifyr

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/spotifyr?color=green)](https://cran.r-project.org/package=spotifyr)
![](http://cranlogs.r-pkg.org/badges/grand-total/spotifyr?color=green)
![](http://cranlogs.r-pkg.org/badges/spotifyr?color=green)
![](http://cranlogs.r-pkg.org/badges/last-week/spotifyr?color=green)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

## Overview
spotifyr is a wrapper for pulling track audio features and other information from Spotify's Web API in bulk. By automatically batching API requests, it allows you to enter an artist's name and retrieve their entire discography in seconds, along with Spotify's audio features and track/album popularity metrics. You can also pull song and playlist information for a given Spotify User (including yourself!).

## Installation
Stable version 1.0.0 on CRAN
```r
install.packages('spotifyr')
```
Development version
```r
devtools::install_github('charlie86/spotifyr')
```

## Authenication
You'll have to set up a Dev account with Spotify to access their Web API [here](https://developer.spotify.com/my-applications/#!/applications). This will give you your `Client ID` and `Client Secret`. Once you have those, you can pull your access token into R with `get_spotify_access_token`. 

The easiest way to authenticate is to set your credentials to the System Environment variables `SPOTIFY_CLIENT_ID` and `SPOTIFY_CLIENT_SECRET`. The default arguments to `get_spotify_access_token` (and all other functions in this package) will refer to those. Alternatively, you can set them manually and make sure to explicitly refer to your access token in each subsequent function call.

```{r, eval = FALSE}
Sys.setenv(SPOTIFY_CLIENT_ID = 'xxxxxxxxxxxxxxxxxxxxx')
Sys.setenv(SPOTIFY_CLIENT_SECRET = 'xxxxxxxxxxxxxxxxxxxxx')

access_token <- get_spotify_access_token()
```

## Usage
### What was The Beatles' favorite key?
```{r load_packages, message = F, warning = F}
library(spotifyr)
```

```{r, message = FALSE, warning = FALSE}
beatles <- get_artist_audio_features('the beatles')

library(tidyverse)
count(beatles, key_mode, sort = T)
```

### What's the most joyful Joy Division song?
My favorite audio feature has to be "valence," a measure of musical positivity.
```{r}
joy <- get_artist_audio_features('joy division')

joy %>% 
 arrange(-valence) %>% 
 select(track_name, valence) %>% 
 head(10)
```
Now if only there was some way to plot joy...

### Joyplot of the emotional rollercoasters that are Joy Division's albums
```{r}
library(ggjoy)

ggplot(joy, aes(x = valence, y = album_name)) + 
  geom_joy() + 
  theme_joy() +
  ggtitle("Joyplot of Joy Division's joy distributions", subtitle = paste0("Based on valence pulled from Spotify's Web API with spotifyr"))
```

## Sentify: A Shiny app
This [app](http://rcharlie.net/sentify/), powered by spotifyr, allows you to visualize the energy and valence (musical positivity) of all of Spotify's artists and playlists.

## Dope stuff other people have done with spotifyr
[Sentiment analysis of musical taste: a cross-European comparison](http://paulelvers.com/post/emotionsineuropeanmusic/) - Paul Elvers
[Blue Christmas: A data-driven search for the most depressing Christmas song](https://caitlinhudon.com/2017/12/22/blue-christmas/) - Caitlin Hudon
[KendRick LamaR](https://davidklaing.github.io/kendrick-lamar-data-science/) - David K. Laing
